---
title: "Statistical Inference Course Project"
author: "Mark Wan"
date: "11/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Overview**
***
This report consists of 2 parts. The first part explores the CLT theorem with the exponential distribution. The second part conducts an inferential analysis on the ToothGrowth dataset.

# **Part 1: Simulations**
***
## **1. Show the sample mean and compare it to the theoretical mean of the distribution.**

The following R code produces 1000 sets of 40 values generated from an exponential distribution with
lambda=0.2. The code then generates a histogram of sample means and labels the actual mean against the
theoretical mean for a direct comparison.

```{r distribution}
# Set reproducibility
set.seed(2020)
lambda <- 0.2
mu_theo <- 1/lambda
s <- 1/lambda
n <- 40
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(n,lambda)))
hist(mns,main="Actual vs Theoretical Mean", xlab="Sample Mean",breaks=40)
abline(v=mean(mns),col="red",lwd=3)
abline(v=mu_theo,col="blue",lwd=3,lty=2)
legend("topright", legend=c("Actual Mean", "Theoretical Mean"),
col=c("red", "blue"), lty=1:2, cex=0.8)
```

The graph above shows that the theoretical mean, 5 is a very good estimate of the actual mean 5.0339482.

## **2. Show how variable the sample is (via variance) and compare it to the theoretical variance**
of the distribution.
The code below does a simple calculation to calculate the theoretical variances vs actual variances.
```{r variances}
var_sample <- var(mns)
var_theo <- s^2/n
```

Attribute | Value
----------|-----------
Theoretical Variance | `r var_theo`
Actual Variance | `r var_sample`
Theoretical Standard Deviation | `r sqrt(var_theo)`
Actual Standard Deviation | `r sqrt(var_sample)`

The table above shows again that the theoretical variance and standard deviation are good estimators of the
actual values.

## **3. Show that the distribution is approximately normal.**
The following R code compares the actual sample mean distribution to the normal distribution.
```{r actual vs theoretical distribution}
hist(mns,prob=TRUE,main="Actual vs Theoretical Mean", xlab="Sample Mean",breaks=40)
lines(density(mns), lwd=3, col="red")
x <- seq(min(mns), max(mns), length=2*n)
y <- dnorm(x, mean=1/lambda, sd=sqrt(((1/lambda)/sqrt(n))^2))
lines(x, y, pch=22, col="blue", lwd=2, lty = 2)
legend("topright", legend=c("Actual Density", "Normal Density"),
col=c("red", "blue"), lty=1:2, cex=0.8)
```
The graph above shows sample mean distribution can be very closely approximated wiht a normal distribution.

# **Part 2: Basic Inferential Data Analysis**
***
The following R code loads the ToothGrowth dataset and executes a basic exploratory analysis of the data.
## **1. Load the ToothGrowth data and perform some basic exploratory data analyses**
```{r exploratory analysis}
data("ToothGrowth")
summary(ToothGrowth)
```

```{r exploratory analysis - head}
head(ToothGrowth)
```

```{r exploratory analysis - unique names}
for (name in names(ToothGrowth)){
print(name)
print(unique(ToothGrowth[,name]))
print("")
}
```

```{r factor dose}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

## **2. Provide a basic summary of the data.**
The following code chunk uses ggplot to produce facet plots
of tooth length vs dose for each supplementary delivery, and of tooth length vs supplmentary delivery for
each dose level.
```{r facet boxplots - supp}
library(ggplot2)
library(RColorBrewer)
g <- ggplot(ToothGrowth,aes(x=dose,y=len)) +
  geom_boxplot(fill=brewer.pal(length(unique(ToothGrowth$dose))*
                                 length(unique(ToothGrowth$supp)),"Paired")) + 
  facet_grid(.~supp) + 
  labs(x="Dose",y="Tooth Length",
       title="Tooth Length vs Dose \n for Different Supplementary Delivery") + 
  theme(plot.title = element_text(hjust = 0.5))
g
```

```{r facet boxplots - dose}
h <- ggplot(ToothGrowth,aes(x=supp,y=len)) + 
  geom_boxplot(fill=brewer.pal(6,"Paired")) + 
  facet_grid(.~dose) + 
  labs(x="Supplementary Delivery",y="Tooth Length",
       title="Tooth Length vs Supplementary Delivery \n for Different Dose") + 
  theme(plot.title = element_text(hjust = 0.5))
h
```
The 2 facet plots above show (1) sample averages of tooth length increased with dose for both OJ and VC
supplementary deliveries, and (2) sample average of tooth length decreased with a change in supplementary
delivery from OJ to VC for dose level 0.5 and 1. However, the sample average between the 2 deliveries did
not differ much for dose level 2.
These observations are only descriptions of the data. In order to make inferences about the relationship
between tooth length and the 2 variables, hypothesis testing is required.

## **3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there’s other approaches worth considering)**
The following 3 code chunks conduct unpaired t-tests for each possible pair of dose levels.

```{r Test Dose 0.5 vs 1}                                                   
# Test Dose 0.5 vs 1
dose_0.5_1 <- subset(ToothGrowth, dose %in% c(0.5,1))
t.test(len~dose,data=dose_0.5_1)
```

```{r Test Dose 0.5 vs 2} 
# Test Dose 0.5 vs 2
dose_0.5_2 <- subset(ToothGrowth, dose %in% c(0.5,2))
t.test(len~dose,data=dose_0.5_2)
```

```{r Test Dose 1 vs 2} 
# Test Dose 1 vs 2
dose_1_2 <- subset(ToothGrowth, dose %in% c(1,2))
t.test(len~dose,data=dose_1_2)
```

```{r Test Supplementary Delivery} 
# Test Supplementary Delivery
t.test(len~supp,data=ToothGrowth)
```
For a 5% significance level, the difference in average between VC and OJ was not significant, since the p-value
was greater than 5%. That is, the data implies that supplmentary delivery has no impact on tooth length.

## **4. State your conclusions and the assumptions needed for your conclusions.**
In conclusion, based on a 5% significance level, the data seems to indicate that tooth length increases
with does level. In contrasts, there were no significant association observed between tooth length and
supplementary delivery.
These observations were based on the assumptions that: 1. The distribution of the sample means follows
the CLT theorem, and can be approximated by the t-distribution, 2. The observations are not biased and
are a true representation of the underlying population of tooth lengths.


